---
title: "STAT 426 Assignment 1"
output: pdf_document
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Due Tuesday, August 31, 11:59 pm.** 

Submit through Moodle.

## Name: Brianna Diaz
### Netid: bdiaz22

Submit your computational work both as an R markdown (*.Rmd) document and as a pdf, along with any files needed to run the code. Embed your answers to each problem in the document below after the question statement. If you have hand-written work, please scan or take pictures of it and include in a pdf file, ideally combined with your pdf output file from R Markdown. Be sure to show your work.


### Problem 1 (10 pts)

An exam has 80 multiple choice questions. Each question has 4 choices for the answer, exactly one of which is correct. Suppose for each question a student guesses one of the answers at random.

**(a)** Specify the distribution of the student's number of correct answers.

**Answer:** This would be a Binomial Distribution. 

**(b)** Find the mean and standard deviation of the distribution in (a). 

**Answer:** 
```{r}
80 * (0.25) -> mean

print(mean)

80 * (0.25) * (0.75) -> a 
sqrt(a) -> SD

print(SD)
```

**(c)** Would it be surprising if the student got at least 40 correct answers? Why?

**Answer:** Yes it would be surprising because the are only expected to get 0.25 right, which is 20 questions correct. 40 correct answers is twice as much than expected.

**(d)** Specify the distribution of $(n_1, n_2, n_3, n_4)$, where $n_j$ is the number of times the student selected choice $j$, for $j=1,2,3,4$.

**Answer:** This would be a multinomial distribution (80, 1/4,1/4,1/4,1/4) 

**(e)** Find $E(n_j)$, $\text{var}(n_j)$, $\text{cov}(n_j, n_k)$ and $\text{corr}(n_j, n_k)$ for $j\ne k$.

**Answer:** 
```{r}
80 -> n
0.25 -> x
0.25 -> y

n*x -> E
print(E)

n*x*(1-x) -> var
print(var)

-(n*x*y) -> cov
print(cov)

n*y*(1-y) -> var2

cov/(sqrt(var*var2)) -> corr
print(corr)

```


### Problem 2 (10 pts)
 
Let $Y_1, Y_2, \ldots, Y_n$ be independent observations from the Poisson distribution with mean $\mu$, which is unknown. Recall that if
$Y = \sum_{i=1}^n Y_i$, then $Y \sim \text{Poisson}(n\mu)$.

**(a)** Under the model assumptions in (a), find mathematical expressions for the likelihood, $\ell(\mu)$ and log-likelihood, $L(\mu)$, simplifying the expressions as much as possible.
**Answer:**

**(b)** Show that the maximum likelihood estimator is $\hat{\mu}=\bar{Y}=Y/n$.

Answer:

**(c)** Find an expression for the exact variance of $\hat{\mu}$.

Answer:

**(d)** Using Fisher Information, find the large sample (asymptotic) variance of $\hat{\mu}$. Is it the same as the variance in (c)?

Answer:

**(e)** Suppose you obtain a sample (3, 6, 3, 3, 5, 4, 6, 9, 3, 3) from the Poisson distribution with mean $\mu$. Find the p-value for a two-sided Wald test of the null hypothesis $H_0: \mu=3$.

Answer:



