---
title: "STAT 426 Assignment 3"
output:
  pdf_document: default
  html_document:
    df_print: paged
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Due Tuesday, September 14, 11:59 pm.** 

Submit through Moodle.

## Name: Brianna Diaz
### Netid: bdiaz22

Submit your work both as an R markdown (*.Rmd) document and as a pdf, along with any files needed to run the code. Embed your answers to each problem in the document below after the question statement. If you have hand-written work, please scan or take pictures of it and include in a pdf file, ideally combined with your pdf output file from R Markdown. Be sure to show your work.


### Problem 1 (6 pts)
 
The relationship between coffee drinking and myocardial infarction (MI) was investigated in a retrospective study of women aged 30-49. The study included cases hospitalized for the occurrence of MI, and controls hospitalized for some other acute condition. Data for coffee consumption versus MI/control status, stratified by smoking status, are given below:

\begin{center}
    \begin{tabular}{lccc}
    & & \multicolumn{1}{c}{Nonsmokers} \\
     Coffee consumption & & \multicolumn{1}{c}{MI} &
       \multicolumn{1}{c}{Controls} \\[0.5ex] \hline \\[-2ex]
    $\ge 5$ cups/day  & & 14 & 49 \\
    $< 5$ cups/day & & 75 & 381 \\[0.5ex] \hline
    \\
    \\
    & & \multicolumn{1}{c}{Smokers} \\
     Coffee consumption & & \multicolumn{1}{c}{MI} &
       \multicolumn{1}{c}{Controls} \\[0.5ex] \hline \\[-2ex]
    $\ge 5$ cups/day  & & 138 & 134 \\
    $< 5$ cups/day & & 260 & 416 \\[0.5ex] \hline
    \end{tabular}
    \end{center}
    
**(a)** Compute the estimated conditional odds ratios for nonsmokers and for smokers. 

Answer:
```{r}
coffee <- data.frame(Drink=c("5 cups or more", "5 cups or more","less then 5 cups", "less then 5 cups"),
Status=c("Nonsmoking", "Smoking", "Nonsmoking", "Smoking"),
MI=c(14,138,75,260),
Controls=c(49,134,381,416)
)
#coffee

c.tab <- xtabs(MI ~ Drink + Status, data=coffee)
#c.tab

#Odds Ratio 
(OR.est <- c.tab[1,1]*c.tab[2,2]/(c.tab[1,2]*c.tab[2,1]))

```

**(b)** Compute the estimated marginal odds ratio for the $2\times 2$ table obtained by summing over smoking status.

Answer:
```{r}
more.coffee <- 14+138
less.coffee <- 75+260
controlm <- 49+134
controll <- 381+416

odds.ratio <- more.coffee*controll/(less.coffee*controlm)
odds.ratio
```

**(c)** Based on the results in (a) and (b), comment on whether stratifying on smoking status was important in this study.

Answer: Yes, stratifying on smoking status was important in this study.

### Problem 2 (8 pts)

This problem refers to the data in Problem 1.

**(a)** Let $\theta_{CM(1)}$ denote the (conditional) odds ratio for coffee consumption (C) versus MI status (M) among nonsmokers only. Compute a 95\% confidence interval for $\theta_{CM(1)}$. 

Answer: 
```{r}
(OR.est.non <- 14*381/(49*75))

logOR.se = sqrt(sum(1/14,1/49,1/75,1/381))
logOR.CI <- log(OR.est.non) + c(-1,1) * qnorm(1-0.05/2) * logOR.se
exp(logOR.CI)

```

**(b)** Interpret the result in (a), explaining whether or not there is a significant association between coffee consumption and MI for nonsmokers, at a significance level $\alpha=0.05$. 

Answer: Based off the results there is an association at the significance level at 0.05. 

**(c)** Let $\theta_{CM(2)}$ denote the (conditional) odds ratio for coffee consumption versus MI status among smokers only. Compute a 95% confidence interval for $\theta_{CM(2)}$.

Answer:
```{r}
(OR.est.smoke <- 138*416/(134*260))

logOR.se.smoke = sqrt(sum(1/138,1/416,1/134,1/260))
logOR.CI.smoke <- log(OR.est.smoke) + c(-1,1) * qnorm(1-0.05/2) * logOR.se.smoke
exp(logOR.CI.smoke)
```

**(d)** Interpret the result in (c), explaining whether or not there is a significant association between coffee consumption and MI for smokers, at a significance level $\alpha=0.05$. 

Answer: Based off the results there is no association at the significance. 

### Problem 3 (6 pts)

This problem again refers to the data in Problem 1 and uses the notation of Problem 2.

**(a)** Using only the \textbf{nonsmoker} data, compute Pearson's $X^2$, the likelihood ratio $G^2$, and their p-values for testing $H_{01}: \theta_{CM(1)}=1$ versus $H_{A1}: \theta_{CM(1)}\ne 1$. In other words test for independence between coffee consumption and MI status conditional on being in the nonsmoker group. What do you conclude? Is there significant association at level .05?

Answer:
```{r}
nonsmoke <- data.frame(nonsmoke=c(14,75), Ctrln=c(49,381))
cn.X2 <- chisq.test(nonsmoke, correct = FALSE)
cn.X2 

cn.G2 <- 2*sum(cn.X2$observed * log(cn.X2$observed/cn.X2$expected))
data.frame(G2=cn.G2, df=1, pvalue=1-pchisq(cn.G2,1))

# There is no significant association at the level 0.05. 
```

**(b)** Using only the \textbf{smoker} data, compute Pearson's $X^2$, the likelihood ratio $G^2$, and their p-values for testing $H_{02}: \theta_{CM(2)}=1$ versus $H_{A2}: \theta_{CM(2)}\ne 1$. In other words test for independence between coffee consumption and MI status conditional on being in the smoker group. What do you conclude? Is there significant association at level .05?

Answer:
```{r}
smoke <- data.frame(smoke=c(138,260), Ctrls=c(134,416))
cs.X2 <- chisq.test(smoke, correct = FALSE)
cs.X2 

cs.G2 <- 2*sum(cs.X2$observed * log(cs.X2$observed/cs.X2$expected))
data.frame(G2=cs.G2, df=1, pvalue=1-pchisq(cs.G2,1))

# There is a significant association at the 0.05. 
```

**(c)** Finally, let's test the null hypothesis of conditional independence of coffee consumption and MI status given smoking status based on all of the data. This is a test of 
$$H_0: \theta_{CM(1)}=1 \text{ and } \theta_{CM(2)}=1$$
versus
$$H_A: \theta_{CM(1)}\ne 1 \text{ or } \theta_{CM(2)}\ne 1.$$ 
Let $G_{tot}^2 = G_1^2 + G_2^2$ where $G_1^2$ is the likelihood ratio $G^2$ statistic from (a) and $G_2^2$ is the likelihood ratio $G^2$ statistic from (b). What is the asymptotic distribution of $G_{tot}^2$ under $H_0$ (including degrees of freedom)? Compute $G_{tot}^2$ and its p-value.

Answer: 

$G_{tot}^2$ is 13.132113 with a p-value of 0.001407321 with a degrees of freedom of 2. 

```{r}
g2.sum <- cs.G2 + cn.G2
g2.sum

no = (nrow(nonsmoke)-1)*(ncol(nonsmoke)-1)
yes = (nrow(smoke)-1)*(ncol(smoke)-1)

df = no + yes
df

1 - pchisq(g2.sum, df = 2)
```

