---
title: "STAT 426 Assignment 7"
output:
  pdf_document: default
  html_document:
    df_print: paged
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Due Tuesday, October 19, 11:59 pm.** 

Submit through Moodle.

## Name: Brianna Diaz
### Netid: bdiaz22

Submit your work both as an R markdown (*.Rmd) document and as a pdf, along with any files needed to run the code. Embed your answers to each problem in the document below after the question statement. If you have hand-written work, please scan or take pictures of it and include in a pdf file, ideally combined with your pdf output file from R Markdown. Be sure to show your work.

### Problem 1. (8 pts) 

The comma-separated data file 'surgery.csv' is included with this assignment. The data concern patients having surgery under general anesthesia. The variables are

Y = whether a patient experienced a sore throat on waking (0 = no, 1 = yes)

D = duration of surgery (in minutes)

T = type of device used to secure the airway (0 = laryngeal mask airway, 1 = tracheal tube)

**a)** (2 pts) Read the data into a data frame and display the first few rows of the data set. (Hint: check R help for 'read.csv')
```{r}
surgery = read.csv("surgery.csv", header = TRUE)
head(surgery)
```

**b)** (2 pts) Fit a logit model that includes \texttt{D}, \texttt{T}, and their interaction. Based on the model summary, is the interaction term significant at level $\alpha=0.05$?
```{r}
logit = glm(Y ~ D * T, family = binomial, data = surgery)

summary(logit)
#p-value: 0.1966, interaction not significant. 
```

**c)** (2 pts) Fit the additive model and display the model summary. Provide likelihood ratio confidence intervals for the coefficients and interpret the effects of the two variables.
```{r}
add_mod = glm(Y ~ D + T, family = binomial, data = surgery)

summary(add_mod)

confint(add_mod)
#The effect the two variables have according to the confidence interval D is significant and T is not 
#This is because D is to the right of ) and T wraps around. 
```

**d)** (2 pts) For the additive model, make a scatter plot of the observed responses versus duration of surgery. Indicate which device was used for each point using the plotting character. (One method is to use pch=T in the plot command). The add two curves to the plot, using different types of dashes or colors: the fitted probability response curve when T=0, and the fitted response curve when T=1. Also include a legend to indicate which curve is which.
```{r}
plot(Y~D, data = surgery, main = "Observed Vs. Durantion", xlab = "Duraton of surgery", ylab= "Observed responses")
curve(predict(add_mod, data.frame(T = 0, D=x), type="response"),col="Green",
add=TRUE)
curve(predict(add_mod, data.frame(T = 1, D=x), type="response"),col="Red",
add=TRUE)
legend("bottomright",
       c("T = 0", "T=1"),
       col = c("Green", "Red"), lwd = 1:2)
```

### Problem 2. (12 pts)

**Install faraway package:** For this exercise, first make sure you have the \texttt{faraway} package installed. Check your R Studio list of packages. If the list includes \texttt{faraway}, you have it. If not, click the 'install' button in the packages window, and start typing \texttt{faraway} into the packages dialog box. It should auto-complete. Select 'faraway' and hit the **install** button.

**Data:** Aflatoxin B1, a type of mold that grows on peanuts and grains, was fed to lab animals at varying doses and the number responding with liver cancer were recorded. \texttt{dose} is the dosage in parts per billion, and, for each dose, \texttt{total} is the number of test animals and \texttt{tumor} is the number with lever cancer. The data are displayed below.

```{r}
library(faraway)
aflatoxin
```

**a)** (2 pts) Consider three link function models for these data:
\begin{align*}
\text{Logit:} &\quad \log\left(\frac{\pi}{1-\pi}\right) = \alpha + \beta\, dose\\
\text{Probit:} &\quad \Phi^{-1}(\pi) = \alpha + \beta\, dose\\
\text{Complementary log-log:} &\quad \log(-\log(1-\pi) = \alpha + \beta\, dose
\end{align*}

Make a scatter plot of the observed proportions with liver cancer versus dose. Add the fitted response curves for the three models to the graph.
```{r}
mod.logit = glm(cbind(tumor,total-tumor) ~ dose,
family=binomial, data=aflatoxin)

mod.probit = glm(cbind(tumor,total-tumor) ~ dose,
family=binomial(link = probit), data=aflatoxin)

mod.cloglog = glm(cbind(tumor,total-tumor) ~ dose,
family=binomial(link = cloglog), data=aflatoxin)

plot(tumor/total~ dose, data=aflatoxin, pch=10, cex=2, cex.axis=1.2, las=1)

curve(predict(mod.logit, data.frame(dose=x), type="response"), add=TRUE, lty=1, col="Blue")

curve(predict(mod.probit, data.frame(dose=x), type="response"), add=TRUE, lty=2, col="Orange", lwd=2)

curve(predict(mod.cloglog, data.frame(dose=x),
type="response"), add=TRUE, lty=3, col="Purple", lwd=3)


```

**b)** (2 pts) Compare the deviances for the three models. Based on the curves and the deviances, which model appears to fit the best?
```{r}
1-pchisq(deviance(mod.logit),df.residual(mod.logit))
1-pchisq(deviance(mod.probit),df.residual(mod.probit))
1-pchisq(deviance(mod.cloglog),df.residual(mod.cloglog))
#The model that appears the best is probit link
```

**c)** (2 pts) Make another scatter plot with three link function curves like the one in a), but this time use $\log(1+dose)$ as the predictor variable, instead of $dose$.
```{r}
mod.logit_log = glm(cbind(tumor,total-tumor) ~ log(1+dose),
family=binomial, data=aflatoxin)

mod.probi_logt = glm(cbind(tumor,total-tumor) ~ log(1+dose),
family=binomial(link = probit), data=aflatoxin)

mod.cloglog_log = glm(cbind(tumor,total-tumor) ~ log(1+dose),
family=binomial(link = cloglog), data=aflatoxin)

plot(tumor/total~ log(1+dose), data=aflatoxin, pch=10, cex=2, cex.axis=1.2, las=1)

curve(predict(mod.logit_log, data.frame(dose = x), type="response"), add=TRUE, lty=1, col="Blue")

curve(predict(mod.probi_logt, data.frame(dose = x), type="response"), add=TRUE, lty=2, col="Orange", lwd=2)

curve(predict(mod.cloglog_log, data.frame(dose = x),
type="response"), add=TRUE, lty=3, col="Purple", lwd=3)

```

**d)** (2 pts) Compare the deviances for the three $log(1+dose)$ models. Based on the curves and deviances, which model appears to fit the best.
```{r}
1 - pchisq(deviance(mod.logit_log), df.residual(mod.logit_log))
1 - pchisq(deviance(mod.probi_logt), df.residual(mod.probi_logt))
1 - pchisq(deviance(mod.cloglog_log), df.residual(mod.cloglog_log))
#The best model is log-log link
```

**e)** (2 pts) Among the 6 different models that you fit, choose which one fits the best and use it to predict the proportion developing liver cancer at a dose of 20.
```{r}
#summary(mod.probit)
#sumary(mod.probit)$coefficients

mod_probit_pre = predict(mod.probit,newdata=data.frame(dose=20), interval="confidence", se.fit=TRUE, type="link")

mod_probit_pre
```

**f)** (2 pts) Provide a 95\% confidence interval for your prediction in **e)**.
```{r}
beta = coefficients(mod.probit)
beta[1] + beta[2]*20

low = -0.6975651 - 1.96*0.1673653

high = -0.6975651 + 1.96*0.167365

c(low, high)
```
